Fast and adaptive cointegration based model for forecasting high frequency financial time series

### 2016-04-22 articles-conferences/iscef2016/src/methodology.tex
comentarios articulo Peretti

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors do not change in time.
In fact, the long-run relationships between the time series might change due to
several economic factors that can lead to structural breaks in the cointegration
relationship \cite{gregoryETal1996}. 
In order to show the the number of cointegration vectors which depends on 
the amount $L$ of historical data and the number of lags $p$ in the VECM, we used a grid search. We defined a grid of possible values for $L$ and $p$. $L$ goes throughout $[100,5000]$ with a step size of 300 and $p$  throughout $[1,5]$ with step size of 1. The first experiment consisted in determining the number of cointegration vectors for all combinations of $L$ and $p$.
We used four forex rates: EURUSD, GBPUSD, USDCHF and USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of August 2014,
when the New York and London financial markets opened.

@@

On the other hand, ARIMA is widely used to forecast returns in finance \cite{tsay2005}. A process can be modelled as an ARIMA$(p,d,q)$ model if $\mathbf{x}_t=\Delta^d \mathbf{y}_t $, i.e after differencing $d$ times the time series $\mathbf{y}_t$,  we get an ARMA$(p,q)$. Since we are modelling returns, we used $d=1$. An ARMA$(p,q)$ model is the following:

\begin{equation}
\mathbf{x}_t = \sum_{i=1}^p \phi_i \mathbf{x}_{t-i}  +  \sum_{j=1}^q \theta_j \epsilon_{t-j}  
\end{equation}
\noindent with coefficients $\phi_p \neq 0$, $\theta_q \neq 0$ and $\sigma_{\epsilon}^2 > 0$.

>>>>

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors do not change in time.
In fact, the long-run relationships between the time series might change due to
several economic factors that can lead to structural breaks in the cointegration
relationship \cite{gregoryETal1996}. 
In order to show the number of cointegration vectors which depends on 
the amount $L$ of historical data and the number of lags $p$ in the VECM, we used a grid search. We defined a grid of possible values for $L$ and $p$. $L$ goes throughout $[100,5000]$ with a step size of 300 and $p$  throughout $[1,5]$ with step size of 1. The first experiment consisted in determining the number of cointegration vectors for all combinations of $L$ and $p$.
We used four forex rates: EURUSD, GBPUSD, USDCHF and USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of August 2014,
when the New York and London financial markets opened.

@@

On the other hand, ARIMA is widely used to forecast returns in finance \cite{tsay2005}. A process can be modelled as an ARIMA$(p,d,q)$ model if $\mathbf{x}_t=\Delta^d \mathbf{y}_t $, i.e after differencing $d$ times the time series $\mathbf{y}_t$,  we get an ARMA$(p,q)$. Since we are modelling returns, we used $d=1$.

### 2016-04-22 articles-conferences/iscef2016/src/results.tex
comentarios articulo Peretti

VECM can only be used if the time series are I(1). Therefore, 
before running the tests, we firstly checked whether the time series were
I(1) using the Augmented Dickey Fuller (ADF) test with lags $p=1,2,3,4,5$. 
MacKinnon \cite{mackinnon2010} critical values for rejection of hypothesis of a unit root are: -2.62 (1\%), 1.94 (5\%) and 1.62 (10\%).
\begin{table*}[ht]
\label{tab:adf}
\centering
\begin{tabular}{llllll}
\toprule
{Variable} & {ADF(1)} & {ADF(2)} & {ADF(3)} & {ADF(4)} & {ADF(5)}\\ 
\midrule
EURUSD &  -0.052   & -0.054  & -0.054  & -0.054  & -0.054  \\
GBPUSD &  -0.744  & -0.784  & -0.805  & -0.837  & -0.846  \\
USDCHF &  -0.476   & -0.493  & -0.493  & -0.495  & -0.502  \\
USDJPY &  0.357   & 0.360  & 0.360  & 0.367  & 0.367  \\
$\Delta$ EURUSD & -128.4*  & -128.4*  & -96.85* & -89.12*   & -89.12*\\
$\Delta$ GBPUSD & -131.4*  & -112.7*  & -102.5* & -92.86*   & -88.29*\\
$\Delta$ USDCHF & -127.8*  & -127.8*  & -96.94* & -88.82*   & -80.79*\\
$\Delta$ USDJPY & -135.1*  & -135.1*  & -101.2* & -101.2*   & -101.2*\\
\bottomrule
\addlinespace[1ex]
\multicolumn{6}{l}{ \textsuperscript{*} Indicates significance at 1\% level} \\
\multicolumn{6}{l}{ \textsuperscript{**} Indicates significance at 5\% level} \\
\multicolumn{6}{l}{ \textsuperscript{***} Indicates significance at 10\% level} \\
\multicolumn{6}{l}{MacKinnon critical values for rejection of hypothesis of a unit root are:}\\
\multicolumn{6}{l}{ -2.62 (1\%), -1.94 (5\%) and -1.62 (10\%)}\\
\multicolumn{6}{l}{ADF($d$) Augmented Dickey-Fuller test with lag $d$} 
\end{tabular}
\caption{Unit roots tests for EURUSD, GBPUSD, USDCHF and USDJPY at 10-seconds
frequency.}
\end{table*}
Table~\ref{tab:adf} shows that all currency rates cannot reject the unit root test in their level form considering different lags, but they rejected it with their first differences. This means that all of
them are I(1) time series and we are allowed to use VECM and therefore our proposed AVECM.

>>>>

Before running the tests, we firstly checked whether the time series were
I(1) using the Augmented Dickey Fuller (ADF) test with lags $p=1,2,3,4,5$. 
MacKinnon \cite{mackinnon2010} critical values for rejection of hypothesis of a unit root are: -2.62 (1\%), 1.94 (5\%) and 1.62 (10\%).
\begin{table*}[ht]
\label{tab:adf}
\centering
\begin{tabular}{llllll}
\toprule
{Variable} & {ADF(1)} & {ADF(2)} & {ADF(3)} & {ADF(4)} & {ADF(5)}\\ 
\midrule
EURUSD &  -0.052   & -0.054  & -0.054  & -0.054  & -0.054  \\
GBPUSD &  -0.744  & -0.784  & -0.805  & -0.837  & -0.846  \\
USDCHF &  -0.476   & -0.493  & -0.493  & -0.495  & -0.502  \\
USDJPY &  0.357   & 0.360  & 0.360  & 0.367  & 0.367  \\
$\Delta$ EURUSD & -128.4*  & -128.4*  & -96.85* & -89.12*   & -89.12*\\
$\Delta$ GBPUSD & -131.4*  & -112.7*  & -102.5* & -92.86*   & -88.29*\\
$\Delta$ USDCHF & -127.8*  & -127.8*  & -96.94* & -88.82*   & -80.79*\\
$\Delta$ USDJPY & -135.1*  & -135.1*  & -101.2* & -101.2*   & -101.2*\\
\bottomrule
\addlinespace[1ex]
\multicolumn{6}{l}{ \textsuperscript{*} Indicates significance at 1\% level} \\
\multicolumn{6}{l}{ \textsuperscript{**} Indicates significance at 5\% level} \\
\multicolumn{6}{l}{ \textsuperscript{***} Indicates significance at 10\% level} \\
\multicolumn{6}{l}{MacKinnon critical values for rejection of hypothesis of a unit root are:}\\
\multicolumn{6}{l}{ -2.62 (1\%), -1.94 (5\%) and -1.62 (10\%)}\\
\multicolumn{6}{l}{ADF($d$) Augmented Dickey-Fuller test with lag $d$} 
\end{tabular}
\caption{Unit roots tests for EURUSD, GBPUSD, USDCHF and USDJPY at 10-seconds
frequency.}
\end{table*}
Table~\ref{tab:adf} shows that all currency rates cannot reject the unit root test in their level form considering different lags, but they rejected it with their first differences. This means that all of
them are I(1) time series and we are allowed to use VECM and therefore our proposed AVECM.
